Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2022-08-23 Paper/Paper.tex
Updates from Overleaf

\begin{abstract}
    Technological innovation has affected workers and labor markets for decades, yet the effects of modern technologies in some settings are still not understood. I add to the understanding of technology implementation's affect on highly technical jobs by investigating changes in physician behavior as a result of a major technology shift in U.S. health care, electronic health records (EHRs). I treat EHR implementation in hospitals as an exogenous treatment to hospitalists and estimate average group time treatment effects on various labor market outcomes. I find that physicians are 10\% more likely to retire, and 4\% more likely to work in an office if they choose not to retire. Finally, for hospitalists who remain exposed to the technology, number of patients seen increases. These results point to adverse effects that can occur when technology is poorly implemented in technical occupations. 
\end{abstract}

@@

\section{Introduction}

How technological innovation affects labor markets is heavily studied, in part because of the vast differences in technology and tasks present even within industries. Yet, there is still much to be learned about modern technology and its interaction with workers. A recent technology roll-out that altered the U.S. health care system dramatically was the electronic health record (EHR). The implementation and use of EHRs were incentivized by the government, leading to a rapid implementation by hospitals and practices nationwide. Since many physicians working in hospitals were exposed to this technology beyond their control, this is an opportune setting to investigate the relationship between technology implementation and labor market decisions for highly skilled workers.

@@

Physicians, as the primary users of EHRs, play an important role in whether the expected benefits come to fruition. The daily tasks a physician faces change drastically when a new system is put in place. Some physicians reported that when using EHRs they are less satisfied with their job and have higher stress levels. Senior physicians in particular “loathe the cumbersome, time-consuming data entry that comes with using EHRs.” (\cite{CollierBurnout}). The frustration of using a new technology raises the cost of working in settings that it, which may lead physicians on the margin to make behavioral changes such as exiting the labor market altogether or shifting towards lower cost work settings. However, the extent to which this frustration actually imposes a meaningful cost is unknown, making physician response an empirical question. Using a difference in difference research design, I estimate the effect of hospital EHR implementation on individual physician labor market choices.

I form a panel of physicians who work in hospitals, hereafter referred to as hospitalists, which spans from 2009-2017. Using CMS Shared Patient Data, I determine which hospitals each hospitalist is connected to. Then, I use the AHA Hospital Survey to determine which hospitals implement an EHR, and thus connect hospitalists to EHR use. Finally, I use Medicare Data on Provider Practice and Specialty (MD-PPAS) for detailed information on the patients seen by each hospitalist. For all outcomes, the independent variable of interest is a binary treatment variable capturing exposure to an EHR. I estimate group time treatment effects of EHR exposure on the following physician decisions: (1) retirement, measured based on zero or missing billable activity in all future years of the panel, (2) where to physically work, measured by fraction of patients seen in an office setting, and location of the physician office, (3) number of patients seen, and (4) billing activity.

This paper contributes to the literature's understanding of how health information technology generally affects health care. There is a robust set of empirical papers examining the effect of EHRs on patient outcomes and hospital costs, but these studies largely use data prior to the subsidization of meaningful EHR use. Despite a large number of case studies that find generally positive effects (improved patient outcomes and decreased cost) (surveyed in \cite{Buntin2011TheResults}), more recent empirical work has found a mixture of results: that outcomes for median patients do not change as a result of EHRs (\cite{Agha2014TheCare}, \cite{McCullough2016HealthCoordination}, \cite{Meyerhoefer}) while newborns and severe patients experience improvement in health outcomes (\cite{Miller2009}, \cite{Freedman2015}, \cite{McCullough2016HealthCoordination}), and that hospital costs only decrease 6 years after implementation, if at all (\cite{Agha2014TheCare}, \cite{dranove2014trillion}). More relevant to individual physicians' productivity, a number of studies consider the effect of EHRs on productivity in specific settings. One study finds that nursing home productivity increases after adopting health IT (\cite{Hitt2016}), but another finds that physician productivity decreases by 11 percent due to EHRs being adopted in primary-care sites (\cite{Meyerhoefer}).

@@

Early research on the inputs to physician retirement decisions found that the most influential factors are personal and financial matters, and that physicians care about their work environment (specifically in the context of managed care organizations), but do not necessarily retire early because of it (\cite{Bahrami2002}). Contrary to this work, I find that EHR implementation led to a 0.006 ppt increase in the likelihood of a senior physician retiring after implementation, a 20\% increase relative to the average retirement for this age group. Physicians less than 60 were 0.002 ppts more likely to retire after exposure, possibly pointing to career switching since retirement is measured as a lack of Medicare patients seen. These results suggest that EHR implementation may have intensified access to care issues for patients, and that there was a surge of experienced physicians who left the labor force and were no longer available to influence physicians learning under them. For those who did not retire, I find that young physicians were .014 ppts (4.6\% relative to the mean) more likely to see patients in an office setting the same year as EHR exposure.  Finally, I find a modest, positive effect on patient count completely driven by younger physicians, and no effect on billing activity for either age group.

Two key assumptions underlie the main analysis. First, I assume that physicians are constrained to use the technology implemented in their work setting. To test this assumption, I investigate the existence of hospital employees with the purpose of using an EHR on behalf of a physician, which I refer to as data assistants. If data assistants are used to ease EHR use for physicians, the results are at least partially driven by their existence. I find that the majority of hiring of data assistants took place in 2013-2014, a two year delay from the vast EHR implementation. Further, my main results are not sensitive to the exclusion of these years, and estimating the effect of EHR implementation only in hospitals without data assistants yields noisy, but similar, results. Second, I assume the decision for hospitals to implement EHRs is exogenous to individual physicians. That is, a physician does not choose their own EHR exposure. While this assumption is reasonable in many settings, there may be instances where a physician is involved in hospital decision making. For this reason, I also consider a sensitivity analysis in which I limit the sample only to hospitals where vertical integration between physician and the hospital is low, as this indicates a low likelihood that EHR implementation is endogenous. The results of the additional analyses indicate that my main findings are not sensitive to endogeneity concerns stemming from joint EHR decision making among physicians and hospitals.

@@

\subsection{Background on EHRs}

EHRs have been an important feature of health care since the 1980s. Early in the technology's existence, health care professionals perceived the technology as a complement to paper records, primarily deployed by large academic medical centers to improve efficiency in billing and/or scheduling. Physicians did not interact directly with these early-generation EHRs, and thus were not drastically affected by their implementation. As technological innovation made computers more portable, the usability of EHRs increased, creating what is known as the "physician workspace": a computer station for a physician to interface directly with an EHR to record patient updates. Despite usability, physicians kept the view that EHRs were purely complementary to paper due to burdensome data entry. Automation in data entry was non-existent, making it extremely time consuming for the user. Even when automation was developed for particular machines that performed monitoring, the hospital still held responsibility for the accuracy of information and therefore required physicians to manually check each data entry (\cite{evans2016electronic}).

@@

\begin{figure}[ht]
    \centering
    \captionsetup{width=.4\linewidth}
    \caption{Screenshot of EHR System}
    \includegraphics[scale=.4]{Objects/epic-ehr-screenshot.jpg}
    \includegraphics[scale=.115]{Objects/EHRimage2.jpg}
    \label{fig:EPIC}
\end{figure}

@@

CMS Shared Patient data records annual information on count of patients and frequency of billing associated with the same patient across providers within a given time period (30, 60, 90, or 180 days). These data are available from 2009 to 2015. For example, if a primary care physician refers a patient to a specialist, then those two physicians have that particular shared patient in common. The number of shared patients are collected in 30, 90, or 180 day intervals. I focus on the 30 day interval, and I employ this data to detect physicians who work closely with hospitals. I limit the entities by tax-code to only include shared patients for physician-hospital pairs, where the physicians in these pairs match the physicians from MD-PPAS. I am specifically interested in primary care physicians who have a close working relationship with hospitals, who do rounds within at least one hospital. The types of physicians are already limited to primary care, but there still may be office-based primary care physicians in the shared patient data. Therefore, I limit the sample of pairs based on a threshold of same day patients which depends on the number of years the physician appears in the data. I only include physician-hospital pairs who have non-missing data with at least 30 shared patients per year, which gives a threshold of work relationship without excluding physicians who have no claims for a number of years.

@@

Next, I consider whether physicians exhibit behavior that suggests they have shifted setting or location of work. I measure this in the data in three ways. First, I consider the fraction of patients a physician sees in an office-based setting, which is directly available in the data. Second, I construct an indicator variable equal to one if a physician sees a positive fraction of patients in an office-based setting. This binary measure of practice setting offers less variation but captures a more discrete changes in which physicians enter or leave the office-based setting entirely. Third, based on the physician's practice zip code, I form an indicator variable equal to one if the physician's list of zip codes changes from one year to the next. While no single variable directly measures physician movement across hospitals, these three outcomes collectively provide insight into whether physicians desire to shift away from EHRs. Finally, I consider outcome variables which explore inputs to overall efficiency in health care, the number of patients seen and total billing activity.

@@

\begin{figure}[t]
\centering
\captionsetup{width=.45\linewidth}
    \caption{Treatment Variables Over Time}
    \includegraphics[scale=.45]{Objects/sum_stats_year.pdf}
    \label{fig:treatmentgraph}
\end{figure}

@@

I treat EHR implementation as an exogenous treatment variable, and I seek to estimate average treatment effects of EHR exposure on various physician outcomes. In this setting, treatment effects are heterogeneous across physicians who are exposed in different years since the underlying characteristics of hospitals who adopt in different years may be correlated with the decision to adopt. Therefore, to avoid negative weighting issues this causes in a classical two way fixed effects event study specification, I instead estimate average treatment effects for a specific group $g$ at time $t$: 
$$ATT(g,t)=\mathbbm{E}[Y_t(g)-Y_t(0)|G_g=1],$$
where $G_g=1$ for those in group $g$. A group indicates all physicians treated, or first exposed to an EHR, in the same year. To estimate the heterogeneous treatment effects, I employ the doubly robust estimator established in \citeauthor{sant2020doubly} (\citeyear{sant2020doubly}). Other estimators that similarly address the concerns yield similar results, presented in Appendix \ref{app:estimators}. I also present classical two way fixed effects results with physician and year fixed effects in Appendix \ref{app:estimators}. These results yield similar patterns with slightly smaller magnitudes.

Once I estimate the $ATT(g,t)$ for each group and year, I aggregate the estimates over groups to a more familiar event study plot. In some cases, I present the de-aggregated estimates to show that results are not driven by a particular year of treatment.  Further, a weighted average of estimates is taken to provide a single $ATT$ value for each outcome; these values are presented in the notes of each table. Simultaneous confidence bands, calculated using bootstrapped standard errors, are presented.

@@

There are several assumptions necessary to identify the parameters of interest, $ATT(g,t)$. First, I assume that treatment is not reversed once it occurs. That is, once a physician is exposed to an EHR, they cannot be un-exposed. At the hospital level, this assumption is supported by both the institutional details of hospital EHR implementation and the definition of exposure. An EHR is a costly technology and requires a significant amount of collaboration to implement. A hospital does not have incentive to un-implement an EHR once it is implemented. They may add features or switch vendors, but do not reverse EHR use. This institutional detail is supported by Figure \ref{fig:hosp_treat}, which shows EHR status of a 2\% random sample of hospitals over time.\footnote{Note that out of the 81 hospitals included in the random sample, 4 of them show a reversal of treatment status. Upon further inspection, each of these hospitals show such a reversal due to a discrepancy in the coding of whether the hospital uses an EHR partially or fully. Those who use the EHR partially are coded as not fully implementing an EHR since it is not clear whether physicians in the sample would be exposed yet or not. Hospitals with this discretion arre coded as implementing an EHR in the first year that they answered "fully" to the survey question.} Further, since we think of treatment at the physician level as exposure to the technology, this assumption is satisfied when thought of as whether the physician has ever used an EHR. Once they are exposed to the EHR for the first time, that experience cannot be forgotten. However, when considering the outcomes of patient count and billing activity, it could be that a physician moves from a hospital that uses an EHR to a hospital that does not, affecting the number of claims observed in the data. This is why I limit the sample of physicians for these outcomes to those who remain with the same set of hospitals throughout the entire period.

@@

As is usual in a difference-in-differences framework, I assume a version of parallel trends based on not-yet-treated units. I assume that, conditional on years of experience, average outcomes for those treated in group $g$ would have followed a parallel trend as those in groups treated in later periods. This assumption would be violated under external conditions in certain years that may be correlated with labor markets, such as a major recession. The time period I study, 2009-2017, is reassuringly stable. In my analysis, I present p-values for a Wald test of pre-trends. In most cases, there is no evidence of a pre-trend. In cases where a pre-trend could be driving the results, I make no strong claims about the effect of EHR exposure.

In order for this to be a causal estimate of EHR exposure on various labor market outcomes, there are additional institutional assumptions necessary. First, I assume that when a physician works in an electronic record utilizing-hospital, they utilize the electronic health record system fully. That is, there are no ways for the physician to continue practicing without using the EHR. I investigate this assumption in Section \ref{sec:dataass}. Second, I assume that a physician's exposure to EHRs is exogenous. That is, the physician does not influence the hospital's decision to implement an EHR. I investigate this assumption in Section \ref{sec:endogeneity}.

>>>>

\begin{abstract}
    Technological innovation has affected workers and labor markets for decades, yet the effects of modern technologies in some vocations with unique incentives are still not understood. I add to the understanding of technology implementation's affect on highly technical jobs by investigating changes in physician behavior as a result of a major technology shift in U.S. health care, electronic health records (EHRs). I treat EHR implementation in hospitals as an exogenous treatment to physicians within the hospital and estimate average group time treatment effects on various labor market outcomes. I find that physicians are 10\% more likely to retire, and 4\% more likely to work in an office if they choose not to retire. Finally, for physicians who remain exposed to the technology, the number of patients seen increases. However, claims per patient decreases. These results point to adverse responses when technology is poorly implemented in technical occupations, but also indicates potential efficiency gains for those who do not respond adversely. 
\end{abstract}

@@

\section{Introduction}
How technological innovation affects labor markets is heavily studied, in part because of the vast differences in technology and tasks present even within industries. Yet, there is still much to be learned about modern technology and its interaction with workers. A recent technology roll-out that altered the U.S. health care system dramatically was the electronic health record (EHR). The implementation and use of EHRs were incentivized by the government, leading to a rapid implementation by hospitals and practices nationwide. Since many physicians working in hospitals were exposed to this technology beyond their control, this is an opportune setting to investigate the relationship between rapid technology implementation and labor market decisions for highly skilled workers.

@@

Physicians, as the primary users of EHRs, play an important role in whether the expected benefits come to fruition. The daily tasks a physician faces change drastically when a new system is put in place. Some physicians reported that when using EHRs they are less satisfied with their job and have higher stress levels. Senior physicians in particular “loathe the cumbersome, time-consuming data entry that comes with using EHRs.” (\cite{CollierBurnout}). The frustration of using a new technology raises the cost of working in settings that use it, which may lead physicians on the margin to make behavioral changes such as exiting the labor market altogether or shifting towards lower cost work settings. However, the extent to which this frustration actually imposes a meaningful cost is unknown, making physician response an empirical question. Using a difference in difference research design, I estimate the effect of hospital EHR implementation on individual physician labor market choices.

I form a panel of physicians who work in hospitals, hereafter referred to as hospitalists, which spans from 2009-2017. Using CMS Shared Patient Data, I identify hospital-physician pairs that indicate the physician is working within the hospital. Then, I use the AHA Hospital Survey to determine which hospitals implement an EHR, and thus connect hospitalists to EHR use. Finally, I use Medicare Data on Provider Practice and Specialty (MD-PPAS) for detailed information on the patients seen by each hospitalist. For all outcomes, the independent variable of interest is a binary treatment variable capturing exposure to an EHR. I estimate group time treatment effects of EHR exposure on the following physician decisions: (1) retirement, measured based on zero or missing billable activity in all future years of the panel, (2) where to physically work, measured by fraction of patients seen in an office setting and location of the physician office, (3) number of patients seen, and (4) patient and billing activity.

This paper contributes to the literature's understanding of how health information technology generally affects health care. There is robust empirical research examining the effect of EHRs on patient outcomes and hospital costs, but these studies largely use data prior to the subsidization of meaningful EHR use. Despite a large number of case studies that find generally positive effects (improved patient outcomes and decreased cost) (surveyed in \cite{Buntin2011TheResults}), empirical work has found a mixture of results: that outcomes for median patients do not change as a result of EHRs (\cite{Agha2014TheCare}, \cite{McCullough2016HealthCoordination}, \cite{Meyerhoefer}) while newborns and severe patients experience improvement in health outcomes (\cite{Miller2009}, \cite{Freedman2015}, \cite{McCullough2016HealthCoordination}), and that hospital costs only decrease 6 years after implementation, if at all (\cite{Agha2014TheCare}, \cite{dranove2014trillion}). More relevant to individual physicians' productivity, a number of studies consider the effect of EHRs on productivity in specific settings. One study finds that nursing home productivity increases after adopting health IT (\cite{Hitt2016}), but another finds that physician productivity decreases by 11 percent due to EHRs being adopted in primary-care sites (\cite{Meyerhoefer}).

@@

Early research on the inputs to physician retirement decisions found that the most influential factors are personal and financial matters, and that physicians care about their work environment (specifically in the context of managed care organizations), but do not necessarily retire early because of it (\cite{Bahrami2002}). Contrary to this result, I find that EHR implementation led to a 0.006 ppt increase in the likelihood of a senior physician retiring after implementation, a 20\% increase relative to the average retirement for this age group. Physicians less than 60 were 0.002 ppts more likely to retire after exposure, possibly pointing to career switching. These results suggest that EHR implementation may have intensified access to care issues for patients, and that there was a surge of experienced physicians who left the labor force and were no longer available to influence early career physicians. For those who did not retire, I find that young physicians were .014 ppts (4.6\% relative to the mean) more likely to see patients in an office setting the same year as EHR exposure.  Finally, I find a modest, positive effect on patient count completely driven by younger physicians, and a decreasing trend in claims per patient.

Two key assumptions underlie the main analysis. First, I assume that physicians are constrained to use the technology implemented in their work setting. To test this assumption, I investigate the existence of hospital employees with the purpose of using an EHR on behalf of a physician, which I refer to as data assistants. If data assistants are used to ease EHR use for physicians, the results are at least partially driven by their existence. I find that the majority of hiring of data assistants took place in 2013-2014, a two year delay from the vast amount of EHR implementation. Further, my main results are not sensitive to the exclusion of these years, and estimating the effect of EHR implementation only in hospitals without data assistants yields noisy, but similar, results. Second, I assume the decision for hospitals to implement EHRs is exogenous to individual physicians. That is, a physician does not choose their own EHR exposure. While this assumption is reasonable in many settings, there may be instances where a physician is involved in hospital decision making. For this reason, I also consider a sensitivity analysis in which I limit the sample only to hospitals where vertical integration between physician and the hospital is low, as this indicates a low likelihood that EHR implementation is endogenous. The results of the additional analyses indicate that my main findings are not sensitive to endogeneity concerns stemming from joint EHR decision making among physicians and hospitals.

@@

\subsection{Background on EHRs}
EHRs have been an important feature of health care since the 1980s. Early in the technology's existence, health care professionals perceived the technology as a complement to paper records, primarily deployed by large academic medical centers to improve efficiency in billing and/or scheduling. Physicians did not interact directly with these early-generation EHRs, and thus were not drastically affected by their implementation. As technological innovation made computers more portable, the usability of EHRs increased, creating what is known as the "physician workspace": a computer station for a physician to interface directly with an EHR to record patient updates. Despite usability, physicians kept the view that EHRs were purely complementary to paper due to burdensome data entry. Automation in data entry was non-existent, making it extremely time consuming for the user. Even when automation was developed for particular machines that performed monitoring, the hospital still held responsibility for the accuracy of information and therefore required physicians to manually check each data entry (\cite{evans2016electronic}).

@@

\begin{figure}[ht]
    \centering
    \captionsetup{width=.4\linewidth}
    \caption{Screenshot of EHR System}
    \includegraphics[scale=.4]{Objects/epic-ehr-screenshot.jpg}
    \includegraphics[scale=.11]{Objects/EHRimage2.jpg}
    \label{fig:EPIC}
\end{figure}

@@

CMS Shared Patient data records annual information on count of Medicare patients and frequency of billing associated with the same patient across providers within a given time period (30, 60, 90, or 180 days). These data are available from 2009 to 2015. For example, if a primary care physician refers a patient to a specialist, then those two physicians have that particular shared patient in common. The number of shared patients are collected in 30, 90, or 180 day intervals. I focus on the 30 day interval, and I employ this data to detect physicians who work closely with hospitals. I limit the entities by tax-code to only include shared patients for physician-hospital pairs, where the physicians in these pairs match the physicians from MD-PPAS. I am specifically interested in primary care physicians who have a close working relationship with hospitals, who do rounds within at least one hospital. The types of physicians are already limited to primary care, but there still may be office-based primary care physicians in the shared patient data. Therefore, I limit the sample of pairs based on a threshold of same day patients which depends on the number of years the physician appears in the data. I only include physician-hospital pairs who have non-missing data with at least 30 shared patients per year, which gives a threshold of work relationship without excluding physicians who have no claims for a number of years.

@@

Next, I consider whether physicians exhibit behavior that suggests they have shifted setting or location of work. I measure this in the data in three ways. First, I consider the fraction of patients a physician sees in an office-based setting, which is directly available in the data. Second, I construct an indicator variable equal to one if a physician sees a positive fraction of patients in an office-based setting. This binary measure of practice setting offers less variation but captures a more discrete change in which physicians enter or leave the office-based setting entirely. Third, based on the physician's practice zip code, I form an indicator variable equal to one if the physician's list of zip codes changes from one year to the next. While no single variable directly measures physician movement across hospitals, these three outcomes collectively provide insight into whether physicians desire to shift away from or towards EHRs. Finally, I consider outcome variables which explore inputs to overall efficiency in health care, the number of patients seen and claims filed per patient.

@@

\begin{figure}[t]
\centering
\captionsetup{width=.45\linewidth}
    \caption{Treatment Variables Over Time}
    \includegraphics[scale=.5]{Objects/sum_stats_year.pdf}
    \label{fig:treatmentgraph}
\end{figure}

@@

For each outcome variable described in Section \ref{sec:outcome}, I treat EHR implementation as an exogenous treatment variable, and I seek to estimate average treatment effects of EHR exposure. In this setting, treatment effects are heterogeneous across physicians who are exposed in different years since the underlying characteristics of hospitals who adopt in different years may be correlated with the decision to adopt. Therefore, to avoid negative weighting issues this causes in a classical two way fixed effects event study specification, I instead estimate average treatment effects for a specific group $g$ at time $t$: 
$$ATT(g,t)=\mathbbm{E}[Y_t(g)-Y_t(0)|G_g=1],$$
where $G_g=1$ for those in group $g$. A group indicates all physicians treated, or first exposed to an EHR, in the same year. To estimate the heterogeneous treatment effects, I employ the doubly robust estimator established in \citeauthor{sant2020doubly} (\citeyear{sant2020doubly}). Other estimators that similarly address the concerns yield similar results, presented in Appendix \ref{app:estimators}. I also present classical two way fixed effects results with physician and year fixed effects in Appendix \ref{app:estimators}. These results yield similar patterns with slightly smaller magnitudes.

Once I estimate the $ATT(g,t)$ for each group and year, I aggregate the estimates over groups to a more familiar event study plot. For select outcomes, I present the original de-aggregated estimates to show that results are not driven by a particular year of treatment.  Further, a weighted average of estimates is taken to provide a single $ATT$ value for each outcome; these values are presented in the notes of each table. Simultaneous confidence bands, calculated using bootstrapped standard errors, are presented.

@@

There are several assumptions necessary to identify the parameters of interest, $ATT(g,t)$. First, I assume that treatment is not reversed once it occurs. That is, once a physician is exposed to an EHR, they cannot be un-exposed. For physicians that do not switch hospitals, this assumption is supported by both the institutional details of hospital EHR implementation and the data itself. An EHR is a costly technology and requires a significant amount of collaboration to implement. A hospital does not have incentive to un-implement an EHR once it is implemented. They may add features or switch vendors, but do not reverse EHR use. This institutional detail is supported by Figure \ref{fig:hosp_treat}, which shows EHR status of a 2\% random sample of hospitals over time.\footnote{Note that out of the 81 hospitals included in the random sample, 4 of them show a reversal of treatment status. Upon further inspection, each of these hospitals show such a reversal due to a discrepancy in the coding of whether the hospital uses an EHR partially or fully. Those who use the EHR partially are coded as not fully implementing an EHR since it is not clear whether physicians in the sample would be exposed yet or not. Hospitals with this discretion arre coded as implementing an EHR in the first year that they answered "fully" to the survey question.} Further, since we think of treatment at the physician level as exposure to the technology, this assumption is satisfied when thought of as whether the physician has ever used an EHR. Once they are exposed to the EHR for the first time, that experience cannot be forgotten. However, when considering the outcomes of patient count and billing activity, it could be that a physician moves from a hospital that uses an EHR to a hospital that does not, affecting the number of claims observed in the data. This is why I limit the sample of physicians for these outcomes to those who remain with the same set of hospitals throughout the entire period. I discuss this limitation to the data further in Section \ref{sec:patientcount}.

@@

As is usual in a difference-in-differences framework, I assume a version of parallel trends based on not-yet-treated units. I assume that, conditional on years of experience, average outcomes for those treated in group $g$ would have followed a parallel trend as those in groups treated in later periods. This assumption would be violated under external conditions in certain years that may be correlated with labor markets, such as a major recession. The time period I study, 2009-2017, is reassuringly stable. In my analysis, I present p-values for a Wald test of pre-trends. In most cases, there is no evidence of a pre-trend. Further, I investigate pre-trends extensively in Appendix \ref{sec:pretrends}.

To interpret these estimators as the causal effect of EHR exposure on various labor market outcomes, there are additional institutional assumptions necessary. First, I assume that when a physician works in an electronic record utilizing-hospital, they utilize the electronic health record system fully. That is, there are no ways for the physician to continue practicing without using the EHR. I investigate this assumption in Section \ref{sec:dataass}. Second, I assume that a physician's exposure to EHRs is exogenous. That is, the physician does not influence the hospital's decision to implement an EHR. I investigate this assumption in Section \ref{sec:endogeneity}.
